```{r packages}
library(tidyverse)
library(reshape2)
library(ggrepel)
```

# Code block for all reference files

```{r reference files}
summary_file <- read.csv('/Volumes/Calvin/TissueCrosstalk/Figure1 items/correlation_count_summary.csv') #Summary file used for Figure 1B and 1C
main_directory <- '/Volumes/Calvin/TissueCrosstalk' 
```

# Code block for all personal functions

```{r functions}
#Function 1
z_score_normalization <- function(x){
  mu_x <- mean(x)
  sigma_x <- sd(x)
  return((x - mu_x)/sigma_x)
}

#Function 2
score_scatter_plot <- function(condition_directory, 
                          source_tissue, 
                          target_tissue, 
                          highlighted_genes,
                          gene_colors,
                          figure_number, 
                          save_plot = "no",
                          x_offset = 0,
                          y_offset = 0.5){
  Ssec_results <- read.csv(file.path(condition_directory, 'Ssec_results.csv')) %>%
    mutate(Source_gene = gsub("Source_", "", Source_gene))
  
  condition <- trimws(strsplit(basename(condition_directory), "_to_")[[1]])
  
  highlighted_df <- Ssec_results %>%
    filter(Source_gene %in% highlighted_genes)
  
  g <- ggplot(Ssec_results, aes(x = p_Ssec_z, y = -r_Ssec_z)) +
    geom_point(size = 1, 
               aes(color = Source_gene),
               alpha = 0.5) +
    labs(title = paste(source_tissue, "to", target_tissue),
         x = "p score",
         y = "r score",
         color = "Source Gene") + 
    xlim(-2, 4) +
    theme(plot.title = element_text(size = 14, h = 0.5, face = 'bold'),
          axis.text.x = element_text(size = 12, color = 'black'),
          axis.text.y = element_text(size = 12, color = 'black'),
          axis.title = element_text(size = 12, face = 'bold'),
          axis.line = element_line(color = 'black'),
          panel.grid = element_blank(),
          panel.background = element_blank(),
          legend.position = "none") +
    geom_point(data = highlighted_df, 
               aes(color = Source_gene), 
               size = 3,
               alpha = 0.7) + 
    scale_color_manual(values = gene_colors) +
    #geom_text_repel used to orient/space text on ggplot, adjust parameters as needed
    geom_text_repel(data = highlighted_df, 
                    aes(label = Source_gene),
                    size = 5,
                    box.padding = 2.5,
                    point.padding = 0.8,
                    fontface = "bold",
                    segment.color = "black",
                    nudge_x = x_offset,
                    nudge_y = y_offset)
    
  print(g) 
  
  if(save_plot == "yes"){
    save_information <- file.path(main_directory, paste0("PDF_Figure_", figure_number,"_", source_tissue,"_to_",target_tissue,".pdf"))
    print(paste0("Saving ", save_information," to ", main_directory))
    ggsave(save_information, width = 6.5, height = 6.5, g, dpi = 300)
  }
}
```

# Code block for generating heat map

```{r figure 1B, heatmap}
heatmap_data <- summary_file %>%
  mutate(Target = gsub("_secretomics", "", Target),
         Z_score = z_score_normalization(Sig_correlations),
         log_10 = log10(Sig_correlations + 1e-6),
         log_10z = z_score_normalization(log_10)) %>%
  rename(Significant_Correlations = Sig_correlations) %>%
  arrange(desc(Significant_Correlations))

unique_ID <- heatmap_data %>%
  distinct(Source)

unique_identifiers <- rev(unique_ID$Source)

n <- length(unique_identifiers)
square_matrix <- matrix(min(heatmap_data$log_10), nrow = n, ncol = n)
rownames(square_matrix) <- unique_identifiers
colnames(square_matrix) <- unique_identifiers

rownames_equal_colnames <- rownames(square_matrix) %in% colnames(square_matrix)

# Loop through the row names and set the diagonal values as -Inf so that it looks gray
for (i in seq_along(rownames(square_matrix))) {
  if (rownames(square_matrix)[i] %in% colnames(square_matrix)) {
    square_matrix[rownames(square_matrix)[i], rownames(square_matrix)[i]] <- -Inf
  }
}
#

for(i in 1:nrow(heatmap_data)){
  row_name <- heatmap_data$Source[i]
  col_name <- heatmap_data$Target[i]
  count <- heatmap_data$log_10[i]
  
  if(row_name %in% rownames(square_matrix) && col_name %in% colnames(square_matrix)) {
    print(paste(row_name, col_name, count))
    square_matrix[row_name, col_name] <- count  # Assign the count to the matrix
  } else {
    warning(paste("Row or column name not found:", row_name, col_name))
  }
}

for_heatmap <- melt(square_matrix, value.name = "log10")
heatmap_plot <- ggplot(for_heatmap, aes(Var1, Var2, fill = log10)) +
  geom_tile() +
  scale_fill_gradient2(low = "white", 
                       mid = "pink", 
                       high = "purple4", 
                       midpoint = 2.5,
                       limits = c(0,5),
                       oob = scales::squish,
                       na.value = rgb(0,0,0, alpha = 0.3)) +
  labs(title = "Heatmap of Significant Correlations", 
      x = "Target Tissues",
      y = "Source Tissues") +
  theme_minimal() + 
  scale_y_discrete(position="right") +
  theme(plot.title = element_text(size = 12, hjust = 0.5, face = 'bold'),
        axis.title.x = element_text(size = 10, face = 'bold', color = 'black'),
        axis.title.y = element_text(size = 10, face = 'bold', color = 'black'),
        axis.text.x = element_text(size = 9, angle = 90, hjust = 1, color = 'black'),
        axis.text.y = element_text(size = 9, color = 'black'),
        axis.line = element_line(color = "black",),
        legend.position = "left"
       )

print(heatmap_plot)
ggsave('/Volumes/Calvin/TissueCrosstalk/Figure_1B_Heatmap.pdf', plot = heatmap_plot, width = 10 , height = 10, dpi = 300)
```

# Code block for generating box plot

```{r figure 1C, box plot}
boxplot_data <- summary_file %>%
  group_by(Source) %>%
  mutate(IQR = IQR(Sig_correlations)) %>%
  ungroup() %>%
  mutate(Source = reorder(Source, -IQR))

main_plot <- ggplot(boxplot_data, aes(x=Source, y=Sig_correlations)) + 
  geom_boxplot(fill = "lightgreen") +
  theme(
    plot.title = element_text(size = 12, hjust = 0.5, face = 'bold'),
    axis.title.x = element_text(size = 10, face = 'bold'),
    axis.title.y = element_text(size = 10, face = 'bold'),
    axis.text.x = element_text(size = 9, angle = 45, hjust = 1, color = 'black'),
    axis.text.y = element_text(size = 9, angle = 45, hjust = 1, color = 'black'),
    panel.grid = element_blank(),
    panel.background = element_blank(),
    axis.line = element_line(color = "black")) +
  labs(title="Number of Signficant Correlations per Tissue", y = "Significant Correlations", x = "Source Tissues")

inset_data <- boxplot_data %>%
  filter(Source %in% levels(boxplot_data$Source)[50:52])
y_min <- min(inset_data$Sig_correlations, na.rm = TRUE)
y_max <- max(inset_data$Sig_correlations, na.rm = TRUE)

inset_plot <- ggplot(inset_data,
                     aes(x=Source, y=Sig_correlations)) +
  geom_boxplot(fill = "lightblue") +
  coord_cartesian(ylim = c(0, NA)) +
  theme_minimal() + 
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(),
        axis.text.x = element_text(size = 9, angle = 45, hjust = 1, color = 'black'),
        axis.text.y = element_text(size = 9, color = 'black'),
        panel.grid = element_blank(),
        axis.line = element_line(color = "black")
#        axis.text.x = element_blank())
)

box_plot <- main_plot +
  annotation_custom(ggplotGrob(inset_plot),
                    xmin = 35, xmax = 45, ymin = 25000, ymax = 90000)

print(box_plot)

ggsave('/Volumes/Calvin/TissueCrosstalk/Figure1C_BoxPlot.pdf', box_plot, width = 12, height = 8, dpi = 300)
```

# Code block for generating scatter plot of r score against p score

## Key tissue types of interest

-   Adipose - Subcutaneous
-   Adipose - Visceral (Omentum)
-   Heart - Left Ventricle
-   Liver Muscle - Skeletal

## Figure 1D, scoring: Adipose(Subcutaneous) to Heart(Left Ventricle)
```{r figure 1D, scoring: Adipose - Subcutaneous to Heart - Left Ventricle}

condition_directory <- file.path(main_directory, 'Tissue_Pairs', 'Adipose - Subcutaneous_to_Heart - Left Ventricle')
source_tissue <- 'Adipose - Subcutaneous'
target_tissue <- 'Heart - Left Ventricle'

highlighted_genes <- c("ADIPOQ","THPO", "AZGP1") #"LGALS3")
gene_colors <- setNames(c("red", "gold", "brown", "brown2"), highlighted_genes)
figure_number = "1D"

score_scatter_plot(condition_directory, source_tissue, target_tissue, highlighted_genes, gene_colors, figure_number, save_plot = "yes", x_offset = 0.25, y_offset = 0.5)

```

## Figure 1E, scoring: Heart(Left Ventricle) to Adipose(Subcutaneous)
```{r figure 1E, scoring: Heart - Left Ventricle to Adipose - Subcutaneous}

condition_directory <- file.path(main_directory, 'Tissue_Pairs', 'Heart - Left Ventricle_to_Adipose - Subcutaneous') 
source_tissue <- 'Heart - Left Ventricle'
target_tissue <- 'Adipose - Subcutaneous'

highlighted_genes <- c("ADIPOQ","THPO","AZGP1") #"LGALS3")
gene_colors <- setNames(c("red", "gold", "brown", "brown2"), highlighted_genes)
figure_number <- '1E'

score_scatter_plot(condition_directory, source_tissue, target_tissue, highlighted_genes, gene_colors, figure_number, save_plot = "yes", x_offset = 0, y_offset = 0.5)
```

## Figure 1F, scoring: Liver to Heart(Left Ventricle)
```{r figure 1F, scoring: Liver to Heart - Left Ventricle}
condition_directory <- file.path(main_directory, 'Tissue_Pairs', 'Liver_to_Heart - Left Ventricle')
source_tissue <- 'Liver'
target_tissue <- 'Heart - Left Ventricle'


highlighted_genes <- c("AHSG", "F11", "INHBE")
gene_colors <- setNames(c("peru", "brown", "maroon"), highlighted_genes)

figure_number <- '1F'

score_scatter_plot(condition_directory, source_tissue, target_tissue, highlighted_genes, gene_colors, figure_number, save_plot = "yes", x_offset = 0.5, y_offset = 0.5)
```

## Figure 1G, scoring: Heart(Left Ventricle) to Liver 
```{r figure 1G, scoring: Heart - Left Ventricle to Liver}
condition_directory <- file.path(main_directory, 'Tissue_Pairs', 'Heart - Left Ventricle_to_Liver')
source_tissue <- 'Heart - Left Ventricle'
target_tissue <- 'Liver'

highlighted_genes <- c("AHSG", "F11", "INHBE")
gene_colors <- setNames(c("peru", "brown", "maroon"), highlighted_genes)

figure_number <- '1G'

score_scatter_plot(condition_directory, source_tissue, target_tissue, highlighted_genes, gene_colors, figure_number, save_plot = "yes", x_offset = 0.5, y_offset = 0.5)
```
